\RedeclareSectionCommand[beforeskip=-1sp]{chapter}
\KOMAoptions{parskip=half}
\chapter{Acknowledgements}

I want to thank Prof.~Dr.~Dr.~Wolfgang~Rhode for giving me all the
opportunities I could have wished for ever since starting to work on my Bachelor's thesis
in 2014.
I'm also grateful to all members of this working group for lots of deep discussions
and fun hours outside of work.
It made my time here really worthwhile.

During my PHD thesis, most of my work was funded by the DFG through the Collaborative Research Center
876 \enquote{Providing Information by Resource-Constrained Data Analysis}.

I want to especially thank Jens Buß,
who basically left me no other choice but to join this working group,
supervised my Bachelor's and Master's thesis and was my office mate for most
of my time here.

Through all this time, Kai Brügge has been my office mate, pair-programming partner,
travel companion and drinking buddy.
Our shared interests in photo- and typography, programming and fine beers bridged the
gap created by me not supporting the correct football team.
Sometimes, people were really surprised, when only one of us came to a meeting or conference.
For all this, I am truly grateful. 

Andrea Teichmann is the heart and soul of our group,
thank you for all the support over the years.

For all the discussions about programming, software development and life in general I want to thank Dominik Neise.
I learned a lot with and from you.

For enabling me to present my work at the International Cosmic Rays Conference 
in Busan, Korea my thanks go to the DAAD.

For discussions about Cherenkov astronomy, open science, physics, analysis techniques, data formats
and more I want to thank Konrad Bernlöhr, Christoph Deil, Tarek Hassan, Karl Kosack and Abelardo Moralejo.

To Dieter Heck, thank you for promptly answering CORSIKA
support emails although \enquote{I have reduced my
activity in cosmic rays to half a day per week as I am now more than 14 years in
retirement.}

I also owe a deep thanks to everyone involved with PeP et al., but especially to Christophe, Tobi, Henning, Marie, Kevin (all of them), Igor and Ismo.
Igor, Kevin and Ismo started the Toolbox Workshop,
which first introduced me to my programming \enquote{mother language} and which I co-organized in the following years.
The tools taught in this workshop are the foundation of this thesis and opened many doors
for me.
These three also first introduced me to the concepts of Open Source Software and reproducible science
and made me think it would be a good idea to try Arch as my first Linux distribution.

For proof-reading this document, I want thank Alexander, Kai, Lena and  Thorben.

All this would have been impossible without the support of my family.
Especially, I want to thank my father for encouraging my curiosity and skepticism and 
my mother and brother for enduring our dinner table discussions.

Finally, to Lena, thank you for everything. 
